### 2023-09-04 proofs/tex/equiv-lemmas.tex
Grammar update: Γ = set

      By inversion and \ruleref{\ottdruleWFTForallLabel},
      $[[Γ ⊢ ∀pas.iN']]$ is equivalent to $[[Γ, pas ⊢ iN']]$,
      and by \cref{lemma:wf-ctxt-equiv}, it is equivalent to 
      $[[Γ, Ord ({pas} ∩ fv iN') ⊢ iN']]$,
      which, by the induction hypothesis, is equivalent to
      $[[Γ, Ord ({pas} ∩ fv iN') ⊢ [mu]iM']]$.

      Analogously, $[[Γ ⊢ ∀pbs.iM']]$ is equivalent to
      $[[Γ, (Ord {pbs} ∩ fv iM') ⊢ iM']]$.
      By \cref{lemma:wf-subst}, it implies 
      $[[Γ, Ord ({pas} ∩ fv iM') ⊢ [mu]iM']]$. And vice versa, 
      $[[Γ, Ord ({pas} ∩ fv iM') ⊢ [mu]iM']]$ implies  
      $[[Γ, Ord ({pbs} ∩ fv iM') ⊢ [mu-1][mu]iM']]$.

      This way, both $[[Γ ⊢ ∀pas.iN']]$ and $[[Γ ⊢ ∀pbs.iM']]$ are equivalent to
      $[[Γ, Ord ({pas} ∩ fv iN') ⊢ [mu]iM']]$. 
    \item For the cases of the positive types, the proofs are symmetric.
  \end{caseof}
\end{proof}

@@

\begin{lemma}[Normalization preserves substitution signature]
  \label{lemma:norm-subst-sig}
  Suppose that $[[σ]]$ is a substitution, $[[Γ1]]$ and $[[Γ2]]$ are contexts.
  Then $[[Γ2 ⊢ σ : Γ1]]$ implies $[[Γ2 ⊢ nf(σ) : Γ1]]$.
\end{lemma}
\begin{proof}
  Suppose that $[[α± ∊ {Γ1}]]$. 
  Then by \cref{corollary:wf-nf}, $[[Γ2 ⊢ nf([σ]α±)]] = [[ [nf(σ)]α± ]]$ 
  is equivalent to $[[Γ2 ⊢ [σ]α±]]$.

  Suppose that $[[α± ∉ {Γ1}]]$. 
  $[[Γ2 ⊢ σ : Γ1]]$ means that $[[ [σ]α± = α± ]]$, 
  and then $[[ [nf(σ)]α± ]] = [[nf([σ]α±)]] = [[nf(α±)]] = [[α±]]$.
\end{proof}

@@

\begin{corollary}[Normalization preserves well-formedness of substitution]
  \label{corollary:wf-s-nf}
  \hfill \\
   $[[Γ2 ⊢ σ : Γ1]] \iff [[Γ2 ⊢ nf(σ) : Γ1]]$
\end{corollary}
\begin{proof}
  Let us prove the forward direction.
  Suppose that $[[α± ∊ {Γ1}]]$.  Let us show that $[[Γ2 ⊢ [nf(σ)]α±]]$.
  By the definition of substitution normalization,
  $[[ [nf(σ)]α± = nf([σ]α±) ]]$. Then by \cref{corollary:wf-nf-algo},
  to show $[[Γ2 ⊢ nf([σ]α±)]]$, it suffices to show $[[Γ2 ⊢ [σ]α±]]$,
  which holds by the assumption $[[Γ2 ⊢ σ : Γ1]]$.

@@

      Let us construct a subsitution $[[pas ⊢ iPs/pbs : pbs]]$ by
      extending $[[mu]]$ with arbitrary positive types on $[[{pbs} \ fv iM]]$.

@@

\begin{lemma}[Subtyping induced by disjoint substitutions]
  \label{lemma:subt-ind-disj-subst}
  If two disjoint substitutions induce subtyping, they are degenerate (so is the
  subtyping).
  Suppose that $[[Γ ⊢ σ1 : Γ1]]$ and $[[Γ ⊢ σ2 : Γ1]]$,
  where $[[{Γi} ⊆ {Γ}]]$ and $[[{Γ1} ∩ {Γ2}= ∅]]$. Then
  \begin{itemize}
  \item[$-$] assuming $[[Γ ⊢ iN]]$,~
    $[[Γ ⊢ [σ1]iN ≤ [σ2]iN]]$ implies $[[Γ ⊢ σi ≈ id : Ord fv iN]]$
  \item[$+$] assuming $[[Γ ⊢ iP]]$,~
    $[[Γ ⊢ [σ1]iP ≥ [σ2]iP]]$ implies $[[Γ ⊢ σi ≈ id : Ord fv iP]]$
  \end{itemize}
\end{lemma}
\begin{proof}
  Proof by induciton on $[[Γ ⊢ iN]]$ (and mutually on $[[Γ ⊢ iP]]$).
  \begin{caseof}
    \item $[[iN]] = [[α⁻]]$\\
      Then $[[Γ ⊢ [σ1]iN ≤ [σ2]iN]]$ is rewritten as $[[Γ ⊢ [σ1]α⁻ ≤ [σ2]α⁻]]$.
      Let us consider the following cases:
      \begin{caseof}
      \item $[[α⁻ ∉ {Γ1}]]$ and $[[α⁻ ∉ {Γ2}]]$ \label{case:var-not-in-ctxts}\\
        Then $[[Γ ⊢ σi ≈ id : α⁻]]$ holds immediately,
        since $[[ [σi] α⁻]] = [[ [id] α⁻]] = [[α⁻]]$ and
        $[[Γ ⊢ α⁻ ≈ α⁻]]$.
      \item $[[α⁻ ∊ {Γ1}]]$ and $[[α⁻ ∊ {Γ2}]]$\\
        This case is not possible by assumption: $[[{Γ1} ∩ {Γ2}= ∅]]$.
      \item $[[α⁻ ∊ {Γ1}]]$ and $[[α⁻ ∉ {Γ2}]]$\\
        Then we have $[[Γ ⊢ [σ1]α⁻ ≤ α⁻]]$,
        which by \cref{corollary:vars-no-proper-subtypes} means $[[Γ ⊢ [σ1]α⁻ ≈ α⁻]]$,
        and hence, $[[Γ ⊢ σ1 ≈ id : α⁻]]$.

        $[[Γ ⊢ σ2 ≈ id : α⁻]]$ holds since $[[ [σ2]α⁻ ]] = [[α⁻]]$,
        similarly to \cref{case:var-not-in-ctxts}.

      \item $[[α⁻ ∉ {Γ1}]]$ and $[[α⁻ ∊ {Γ2}]]$\\
        Then we have $[[Γ ⊢ α⁻ ≤ [σ2]α⁻]]$,
        which by \cref{corollary:vars-no-proper-subtypes} means $[[Γ ⊢ α⁻ ≈ [σ2]α⁻]]$,
        and hence, $[[Γ ⊢ σ2 ≈ id : α⁻]]$.

        $[[Γ ⊢ σ1 ≈ id : α⁻]]$ holds since $[[ [σ1]α⁻ ]] = [[α⁻]]$,
        similarly to \cref{case:var-not-in-ctxts}.
      \end{caseof}
  \item $[[iN]] = [[∀pas.iM]]$\\
    Then by inversion, $[[Γ, pas ⊢ iM]]$.
    $[[Γ ⊢ [σ1]iN ≤ [σ2]iN]]$ is rewritten as $[[Γ ⊢ [σ1]∀pas.iM ≤ [σ2]∀pas.iM]]$.
    By the congruence of substitution and by the inversion of
    \ruleref{\ottdruleDOneForallLabel}, $[[Γ, pas ⊢ [iQs/pas][σ1]iM ≤ [σ2]iM]]$,
    where $[[Γ, pas ⊢ iQi]]$.
    Let us denote the (Kleisli) composition of $[[σ1]]$ and $[[iQs/pas]]$ as
    $[[σ1']]$, noting that $[[Γ, pas ⊢ σ1' : Γ1, pas]]$,
    and $[[{Γ1, pas} ∩ {Γ2} = ∅]]$.

    Let us apply the induction hypothesis to $[[iM]]$ and the
    substitutions $[[σ1']]$ and $[[σ2]]$ with
    $[[Γ, pas ⊢ [σ1']iM ≤ [σ2]iM]]$ to obtain:
    \begin{align}
      [[Γ, pas ⊢ σ1' ≈ id : Ord fv iM]] \label{fact:subs-proper-sub:forall-ih}\\
      [[Γ, pas ⊢ σ2 ≈ id : Ord fv iM]]  \label{fact:subs-proper-sub:forall-ih2}
    \end{align}

    Then $[[Γ ⊢ σ2 ≈ id : Ord fv ∀pas.iM]]$ holds by strengthening of
    \ref{fact:subs-proper-sub:forall-ih2}:
    for any $[[β±]] \in [[fv ∀pas.iM]] = [[fv iM \ {pas}]]$,
    $[[Γ, pas ⊢ [σ2]β± ≈ β±]]$ is strengthened to $[[Γ ⊢ [σ2]β± ≈ β±]]$, because
    $[[fv [σ2]β±]] = [[fv β±]] = \{[[β±]]\} \subseteq [[{Γ}]]$.

    To show that $[[Γ ⊢ σ1 ≈ id : Ord fv ∀pas.iM]]$, let us take an arbitrary
    $[[β±]] \in [[fv ∀pas.iM]] = [[fv iM \ {pas}]]$.

    $
    \begin{aligned}[t]
      [[β±]] &= [[ [id]β± ]]
             && \text{by definition of $[[id]]$}\\
             &\eqDOne [[ [σ1']β± ]]
             && \text{by \ref{fact:subs-proper-sub:forall-ih}}\\
             &= [[ [iQs/pas][σ1]β±]]
             && \text{by definition of $[[σ1']]$}\\
             &= [[ [σ1]β± ]]
             && \text{because $[[{pas} ∩ fv [σ1]β±]] \subseteq [[{pas} ∩ {Γ}]] = \emptyset$}
    \end{aligned}
    $\\
    This way, $[[Γ ⊢ [σ1]β± ≈ β±]]$ for any $[[β±]] \in [[fv ∀pas.iM]]$ and thus,
    $[[Γ ⊢ σ1 ≈ id : Ord fv ∀pas.iM]]$.

@@

    Applying the induction hypothesis to $[[Γ ⊢ [σ1]iP ≥ [σ2]iP]]$
    and to $[[Γ ⊢ [σ1]iM ≤ [σ2]iM]]$, we obtain (respectively):
    \begin{align}
      &[[Γ ⊢ σi ≈ id : Ord fv iP]] \label{fact:subs-proper-sub:arrow-ih1}\\
      &[[Γ ⊢ σi ≈ id : Ord fv iM]] \label{fact:subs-proper-sub:arrow-ih2}
    \end{align}

    Noting that $[[fv (iP → iM)]] = [[fv iP ∪ fv iM]]$,
    we combine
    \cref{fact:subs-proper-sub:arrow-ih1,fact:subs-proper-sub:arrow-ih2}
    to conclude:
    $[[Γ ⊢ σi ≈ id : Ord fv (iP → iM)]]$.

@@

    Applying the induction hypothesis to $[[Γ ⊢ [σ1]iP ≥ [σ2]iP]]$, we obtain
    $[[Γ ⊢ σi ≈ id : Ord fv iP]]$. Since $[[fv ↑iP]] = [[fv iP]]$, we can
    conclude: $[[Γ ⊢ σi ≈ id : Ord fv ↑iP]]$.
  \item The positive cases are proved symmetrically.
  \end{caseof}
\end{proof}

\begin{corollary}[Substitution cannot induce proper subtypes or supertypes] \label{corollary:subst-proper-subt}
  Assuming all mentioned types are well-formed in $[[Γ]]$ and $[[σ]]$ is a
  substitution $[[Γ ⊢ σ : Γ]]$,
  \begin{align*}
    [[Γ ⊢ [σ]iN ≤ iN]] ~&\Rightarrow~ [[Γ ⊢ [σ]iN ≈ iN]]
                          \text{ and } [[Γ ⊢ σ ≈ id : Ord fv iN]] \\
    [[Γ ⊢ iN ≤ [σ]iN]] ~&\Rightarrow~ [[Γ ⊢ iN ≈ [σ]iN]]
                          \text{ and } [[Γ ⊢ σ ≈ id : Ord fv iN]] \\
    [[Γ ⊢ [σ]iP ≥ iP]] ~&\Rightarrow~ [[Γ ⊢ [σ]iP ≈ iP]]
                          \text{ and } [[Γ ⊢ σ ≈ id : Ord fv iP]] \\
    [[Γ ⊢ iP ≥ [σ]iP]] ~&\Rightarrow~ [[Γ ⊢ iP ≈ [σ]iP]]
                          \text{ and } [[Γ ⊢ σ ≈ id : Ord fv iP]] \\
  \end{align*}
\end{corollary}

\begin{lemma} \label{lemma:mutual-subst-subtyping}
  Asssuming that the mentioned types ($[[iP]]$, $[[iQ]]$, $[[iN]]$, and $[[iM]]$)
  are well-formed in $[[Γ]]$ and that the substitutions ($[[σ1]]$ and $[[σ2]]$) have signature $[[Γ ⊢ σi : Γ]]$,
  \begin{itemize}
  \item[$+$] if $[[Γ ⊢ [σ1] iP ≥ iQ]]$ and $[[Γ ⊢ [σ2] iQ ≥ iP]]$\\
    then there exists a bijection $[[μ : fv iP ↔ fv iQ]]$ such that
    $[[Γ ⊢ σ1 ≈ Sub μ : Ord fv iP]]$ and $[[Γ ⊢ σ2 ≈ Sub μ-1 : Ord fv iQ]]$;
  \item[$-$] if $[[Γ ⊢ [σ1] iN ≤ iM]]$ and $[[Γ ⊢ [σ2] iN ≤ iM]]$\\
    then there exists a bijection $[[μ : fv iN ↔ fv iM]]$ such that
    $[[Γ ⊢ σ1 ≈ Sub μ : Ord fv iN]]$ and $[[Γ ⊢ σ2 ≈ Sub μ-1 : Ord fv iM]]$.
  \end{itemize}
\end{lemma}
\begin{proof}
  \hfill
  \begin{itemize}
  \item[$+$]
    Applying $[[σ2]]$ to both sides of
    $[[Γ ⊢ [σ1] iP ≥ iQ]]$ (by \cref{lemma:subst-pres-subt}),
    we have: $[[Γ ⊢ [σ2 ○ σ1] iP ≥ [σ2]iQ]]$.
    Composing it with $[[Γ ⊢ [σ2] iQ ≥ iP]]$ by transitivity 
    (\cref{lemma:subtyping-transitivity}),
    we have $[[Γ ⊢ [σ2 ○ σ1] iP ≥ iP]]$.
    Then by \cref{corollary:subst-proper-subt},
    $[[Γ ⊢ σ2 ○ σ1 ≈ id : Ord fv iP]]$.
    By a symmetric argument, we also have:
    $[[Γ ⊢ σ1 ○ σ2 ≈ id : Ord fv iQ]]$.

    Now, we prove that
    $[[Γ ⊢ σ2 ○ σ1 ≈ id : Ord fv iP]]$ and
    $[[Γ ⊢ σ1 ○ σ2 ≈ id : Ord fv iQ]]$
    implies that $[[σ1]]$ and $[[σ1]]$
    are (equivalent to) mutually inverse bijections.

@@

\begin{lemma}[Equivalence of polymorphic types]
  \label{lemma:poly-types-equivalence}
  \hfill
  \begin{itemize}
    \item[$-$] For $[[Γ ⊢ ∀pas.iN]]$ and $[[Γ ⊢ ∀pbs.iM]]$,\\ if $[[Γ ⊢ ∀pas.iN ≈ ∀pbs.iM ]]$
    then there exists a bijection $[[μ : {pbs} ∩ fv iM ↔ {pas} ∩ fv iN]]$
    such that $[[ Γ, pas ⊢ iN ≈ [Sub μ] iM ]]$,
    \item[$+$] For $[[Γ ⊢ ∃nas.iP]]$ and $[[Γ ⊢ ∃nbs.iQ]]$,\\  if $[[Γ ⊢ ∃nas.iP ≈ ∃nbs.iQ ]]$
    then there exists a bijection $[[μ : {nbs} ∩ fv iQ ↔ {nas} ∩ fv iP]]$
    such that $[[ Γ, nbs ⊢ iP ≈ [Sub μ] iQ ]]$.
  \end{itemize}
\end{lemma}
\begin{proof}
    \hfill
  \begin{itemize}
    \item[$-$]
      First, by $\alpha$-conversion, we ensure $[[{pas} ∩ fv iM = ∅]]$ and $[[{pbs} ∩ fv iN = ∅]]$.
      By inversion, $[[Γ ⊢ ∀pas.iN ≈ ∀pbs.iM ]]$ implies 
      \begin{enumerate} 
        \item $[[Γ,pbs ⊢ [σ1]iN ≤ iM]]$ for $[[ Γ,pbs ⊢ σ1 : pas ]]$ and 
        \item $[[Γ,pas ⊢ [σ2]iM ≤ iN]]$ for $[[ Γ,pas ⊢ σ2 : pbs ]]$.
      \end{enumerate}
      To apply \cref{lemma:mutual-subst-subtyping}, we weaken 
      and rearrange the contexts, and extend the substitutions to act as identity
      outside of their initial domain:
      \begin{enumerate} 
        \item $[[Γ,pas,pbs ⊢ [σ1]iN ≤ iM]]$ for $[[ Γ,pas,pbs ⊢ σ1 : Γ,pas,pbs ]]$ and 
        \item $[[Γ,pas,pbs ⊢ [σ2]iM ≤ iN]]$ for $[[ Γ,pas,pbs ⊢ σ2 : Γ,pas,pbs ]]$.
      \end{enumerate}
      Then from \cref{lemma:mutual-subst-subtyping}, 
      there exists a bijection $[[μ : fv iM ↔ fv iN]]$ such that 
      $[[Γ,pas,pbs ⊢ σ2 ≈ Sub μ : Ord fv iM]]$ and 
      $[[Γ,pas,pbs ⊢ σ1 ≈ Sub μ-1 : Ord fv iN]]$.

>>>>

      By inversion and \ruleref{\ottdruleWFTForallLabel},
      $[[Γ ⊢ ∀pas.iN']]$ is equivalent to $[[Γ, pas ⊢ iN']]$,
      and by \cref{lemma:wf-ctxt-equiv}, it is equivalent to 
      $[[Γ,  ({pas} ∩ fv iN') ⊢ iN']]$,
      which, by the induction hypothesis, is equivalent to
      $[[Γ,  ({pas} ∩ fv iN') ⊢ [mu]iM']]$.

      Analogously, $[[Γ ⊢ ∀pbs.iM']]$ is equivalent to
      $[[Γ, ( {pbs} ∩ fv iM') ⊢ iM']]$.
      By \cref{lemma:wf-subst}, it implies 
      $[[Γ,  ({pas} ∩ fv iM') ⊢ [mu]iM']]$. And vice versa, 
      $[[Γ,  ({pas} ∩ fv iM') ⊢ [mu]iM']]$ implies  
      $[[Γ,  ({pbs} ∩ fv iM') ⊢ [mu-1][mu]iM']]$.

      This way, both $[[Γ ⊢ ∀pas.iN']]$ and $[[Γ ⊢ ∀pbs.iM']]$ are equivalent to
      $[[Γ,  ({pas} ∩ fv iN') ⊢ [mu]iM']]$. 
    \item For the cases of the positive types, the proofs are symmetric.
  \end{caseof}
\end{proof}

@@

\begin{lemma}[Normalization preserves substitution signature]
  \label{lemma:norm-subst-sig}
  Suppose that $[[σ]]$ is a substitution, $[[Γ1]]$ and $[[Γ2]]$ are contexts.
  Then $[[Γ2 ⊢ σ : Γ1]]$ implies $[[Γ2 ⊢ nf(σ) : Γ1]]$.
\end{lemma}
\begin{proof}
  Suppose that $[[α± ∊ Γ1]]$. 
  Then by \cref{corollary:wf-nf}, $[[Γ2 ⊢ nf([σ]α±)]] = [[ [nf(σ)]α± ]]$ 
  is equivalent to $[[Γ2 ⊢ [σ]α±]]$.

  Suppose that $[[α± ∉ Γ1]]$. 
  $[[Γ2 ⊢ σ : Γ1]]$ means that $[[ [σ]α± = α± ]]$, 
  and then $[[ [nf(σ)]α± ]] = [[nf([σ]α±)]] = [[nf(α±)]] = [[α±]]$.
\end{proof}

@@

\begin{corollary}[Normalization preserves well-formedness of substitution]
  \label{corollary:wf-s-nf}
  \hfill \\
   $[[Γ2 ⊢ σ : Γ1]] \iff [[Γ2 ⊢ nf(σ) : Γ1]]$
\end{corollary}
\begin{proof}
  Let us prove the forward direction.
  Suppose that $[[α± ∊ Γ1]]$.  Let us show that $[[Γ2 ⊢ [nf(σ)]α±]]$.
  By the definition of substitution normalization,
  $[[ [nf(σ)]α± = nf([σ]α±) ]]$. Then by \cref{corollary:wf-nf-algo},
  to show $[[Γ2 ⊢ nf([σ]α±)]]$, it suffices to show $[[Γ2 ⊢ [σ]α±]]$,
  which holds by the assumption $[[Γ2 ⊢ σ : Γ1]]$.

@@

      Let us construct a substitution $[[{pas} ⊢ iPs/pbs : {pbs}]]$ by
      extending $[[mu]]$ with arbitrary positive types on $[[{pbs} \ fv iM]]$.

@@

\begin{lemma}[Subtyping induced by disjoint substitutions]
  \label{lemma:subt-ind-disj-subst}
  If two disjoint substitutions induce subtyping, they are degenerate (so is the
  subtyping).
  Suppose that $[[Γ ⊢ σ1 : Γ1]]$ and $[[Γ ⊢ σ2 : Γ1]]$,
  where $[[Γi ⊆ Γ]]$ and $[[Γ1 ∩ Γ2= ∅]]$. Then
  \begin{itemize}
  \item[$-$] assuming $[[Γ ⊢ iN]]$,~
    $[[Γ ⊢ [σ1]iN ≤ [σ2]iN]]$ implies $[[Γ ⊢ σi ≈ id :  fv iN]]$
  \item[$+$] assuming $[[Γ ⊢ iP]]$,~
    $[[Γ ⊢ [σ1]iP ≥ [σ2]iP]]$ implies $[[Γ ⊢ σi ≈ id :  fv iP]]$
  \end{itemize}
\end{lemma}
\begin{proof}
  Proof by induciton on $[[Γ ⊢ iN]]$ (and mutually on $[[Γ ⊢ iP]]$).
  \begin{caseof}
    \item $[[iN]] = [[α⁻]]$\\
      Then $[[Γ ⊢ [σ1]iN ≤ [σ2]iN]]$ is rewritten as $[[Γ ⊢ [σ1]α⁻ ≤ [σ2]α⁻]]$.
      Let us consider the following cases:
      \begin{caseof}
      \item $[[α⁻ ∉ Γ1]]$ and $[[α⁻ ∉ Γ2]]$ \label{case:var-not-in-ctxts}\\
        Then $[[Γ ⊢ σi ≈ id : {α⁻}]]$ holds immediately,
        since $[[ [σi] α⁻]] = [[ [id] α⁻]] = [[α⁻]]$ and
        $[[Γ ⊢ α⁻ ≈ α⁻]]$.
      \item $[[α⁻ ∊ Γ1]]$ and $[[α⁻ ∊ Γ2]]$\\
        This case is not possible by assumption: $[[Γ1 ∩ Γ2= ∅]]$.
      \item $[[α⁻ ∊ Γ1]]$ and $[[α⁻ ∉ Γ2]]$\\
        Then we have $[[Γ ⊢ [σ1]α⁻ ≤ α⁻]]$,
        which by \cref{corollary:vars-no-proper-subtypes} means $[[Γ ⊢ [σ1]α⁻ ≈ α⁻]]$,
        and hence, $[[Γ ⊢ σ1 ≈ id : {α⁻}]]$.

        $[[Γ ⊢ σ2 ≈ id : {α⁻}]]$ holds since $[[ [σ2]α⁻ ]] = [[α⁻]]$,
        similarly to \cref{case:var-not-in-ctxts}.

      \item $[[α⁻ ∉ Γ1]]$ and $[[α⁻ ∊ Γ2]]$\\
        Then we have $[[Γ ⊢ α⁻ ≤ [σ2]α⁻]]$,
        which by \cref{corollary:vars-no-proper-subtypes} means $[[Γ ⊢ α⁻ ≈ [σ2]α⁻]]$,
        and hence, $[[Γ ⊢ σ2 ≈ id : {α⁻}]]$.

        $[[Γ ⊢ σ1 ≈ id : {α⁻}]]$ holds since $[[ [σ1]α⁻ ]] = [[α⁻]]$,
        similarly to \cref{case:var-not-in-ctxts}.
      \end{caseof}
  \item $[[iN]] = [[∀pas.iM]]$\\
    Then by inversion, $[[Γ, pas ⊢ iM]]$.
    $[[Γ ⊢ [σ1]iN ≤ [σ2]iN]]$ is rewritten as $[[Γ ⊢ [σ1]∀pas.iM ≤ [σ2]∀pas.iM]]$.
    By the congruence of substitution and by the inversion of
    \ruleref{\ottdruleDOneForallLabel}, $[[Γ, pas ⊢ [iQs/pas][σ1]iM ≤ [σ2]iM]]$,
    where $[[Γ, pas ⊢ iQi]]$.
    Let us denote the (Kleisli) composition of $[[σ1]]$ and $[[iQs/pas]]$ as
    $[[σ1']]$, noting that $[[Γ, pas ⊢ σ1' : Γ1, pas]]$,
    and $[[(Γ1, pas) ∩ Γ2 = ∅]]$.

    Let us apply the induction hypothesis to $[[iM]]$ and the
    substitutions $[[σ1']]$ and $[[σ2]]$ with
    $[[Γ, pas ⊢ [σ1']iM ≤ [σ2]iM]]$ to obtain:
    \begin{align}
      [[Γ, pas ⊢ σ1' ≈ id :  fv iM]] \label{fact:subs-proper-sub:forall-ih}\\
      [[Γ, pas ⊢ σ2 ≈ id :  fv iM]]  \label{fact:subs-proper-sub:forall-ih2}
    \end{align}

    Then $[[Γ ⊢ σ2 ≈ id :  fv ∀pas.iM]]$ holds by strengthening of
    \ref{fact:subs-proper-sub:forall-ih2}:
    for any $[[β±]] \in [[fv ∀pas.iM]] = [[fv iM \ {pas}]]$,
    $[[Γ, pas ⊢ [σ2]β± ≈ β±]]$ is strengthened to $[[Γ ⊢ [σ2]β± ≈ β±]]$, because
    $[[fv [σ2]β±]] = [[fv β±]] = \{[[β±]]\} \subseteq [[Γ]]$.

    To show that $[[Γ ⊢ σ1 ≈ id :  fv ∀pas.iM]]$, let us take an arbitrary
    $[[β±]] \in [[fv ∀pas.iM]] = [[fv iM \ {pas}]]$.

    $
    \begin{aligned}[t]
      [[β±]] &= [[ [id]β± ]]
             && \text{by definition of $[[id]]$}\\
             &\eqDOne [[ [σ1']β± ]]
             && \text{by \ref{fact:subs-proper-sub:forall-ih}}\\
             &= [[ [iQs/pas][σ1]β±]]
             && \text{by definition of $[[σ1']]$}\\
             &= [[ [σ1]β± ]]
             && \text{because $[[{pas} ∩ fv [σ1]β±]] \subseteq [[{pas} ∩ Γ]] = \emptyset$}
    \end{aligned}
    $\\
    This way, $[[Γ ⊢ [σ1]β± ≈ β±]]$ for any $[[β±]] \in [[fv ∀pas.iM]]$ and thus,
    $[[Γ ⊢ σ1 ≈ id :  fv ∀pas.iM]]$.

@@

    Applying the induction hypothesis to $[[Γ ⊢ [σ1]iP ≥ [σ2]iP]]$
    and to $[[Γ ⊢ [σ1]iM ≤ [σ2]iM]]$, we obtain (respectively):
    \begin{align}
      &[[Γ ⊢ σi ≈ id :  fv iP]] \label{fact:subs-proper-sub:arrow-ih1}\\
      &[[Γ ⊢ σi ≈ id :  fv iM]] \label{fact:subs-proper-sub:arrow-ih2}
    \end{align}

    Noting that $[[fv (iP → iM)]] = [[fv iP ∪ fv iM]]$,
    we combine
    \cref{fact:subs-proper-sub:arrow-ih1,fact:subs-proper-sub:arrow-ih2}
    to conclude:
    $[[Γ ⊢ σi ≈ id :  fv (iP → iM)]]$.

@@

    Applying the induction hypothesis to $[[Γ ⊢ [σ1]iP ≥ [σ2]iP]]$, we obtain
    $[[Γ ⊢ σi ≈ id :  fv iP]]$. Since $[[fv ↑iP]] = [[fv iP]]$, we can
    conclude: $[[Γ ⊢ σi ≈ id :  fv ↑iP]]$.
  \item The positive cases are proved symmetrically.
  \end{caseof}
\end{proof}

\begin{corollary}[Substitution cannot induce proper subtypes or supertypes] \label{corollary:subst-proper-subt}
  Assuming all mentioned types are well-formed in $[[Γ]]$ and $[[σ]]$ is a
  substitution $[[Γ ⊢ σ : Γ]]$,
  \begin{align*}
    [[Γ ⊢ [σ]iN ≤ iN]] ~&\Rightarrow~ [[Γ ⊢ [σ]iN ≈ iN]]
                          \text{ and } [[Γ ⊢ σ ≈ id :  fv iN]] \\
    [[Γ ⊢ iN ≤ [σ]iN]] ~&\Rightarrow~ [[Γ ⊢ iN ≈ [σ]iN]]
                          \text{ and } [[Γ ⊢ σ ≈ id :  fv iN]] \\
    [[Γ ⊢ [σ]iP ≥ iP]] ~&\Rightarrow~ [[Γ ⊢ [σ]iP ≈ iP]]
                          \text{ and } [[Γ ⊢ σ ≈ id :  fv iP]] \\
    [[Γ ⊢ iP ≥ [σ]iP]] ~&\Rightarrow~ [[Γ ⊢ iP ≈ [σ]iP]]
                          \text{ and } [[Γ ⊢ σ ≈ id :  fv iP]] \\
  \end{align*}
\end{corollary}

\begin{lemma} \label{lemma:mutual-subst-subtyping}
  Asssuming that the mentioned types ($[[iP]]$, $[[iQ]]$, $[[iN]]$, and $[[iM]]$)
  are well-formed in $[[Γ]]$ and that the substitutions ($[[σ1]]$ and $[[σ2]]$) have signature $[[Γ ⊢ σi : Γ]]$,
  \begin{itemize}
  \item[$+$] if $[[Γ ⊢ [σ1] iP ≥ iQ]]$ and $[[Γ ⊢ [σ2] iQ ≥ iP]]$\\
    then there exists a bijection $[[μ : fv iP ↔ fv iQ]]$ such that
    $[[Γ ⊢ σ1 ≈ Sub μ :  fv iP]]$ and $[[Γ ⊢ σ2 ≈ Sub μ-1 :  fv iQ]]$;
  \item[$-$] if $[[Γ ⊢ [σ1] iN ≤ iM]]$ and $[[Γ ⊢ [σ2] iN ≤ iM]]$\\
    then there exists a bijection $[[μ : fv iN ↔ fv iM]]$ such that
    $[[Γ ⊢ σ1 ≈ Sub μ :  fv iN]]$ and $[[Γ ⊢ σ2 ≈ Sub μ-1 :  fv iM]]$.
  \end{itemize}
\end{lemma}
\begin{proof}
  \hfill
  \begin{itemize}
  \item[$+$]
    Applying $[[σ2]]$ to both sides of
    $[[Γ ⊢ [σ1] iP ≥ iQ]]$ (by \cref{lemma:subst-pres-subt}),
    we have: $[[Γ ⊢ [σ2 ○ σ1] iP ≥ [σ2]iQ]]$.
    Composing it with $[[Γ ⊢ [σ2] iQ ≥ iP]]$ by transitivity 
    (\cref{lemma:subtyping-transitivity}),
    we have $[[Γ ⊢ [σ2 ○ σ1] iP ≥ iP]]$.
    Then by \cref{corollary:subst-proper-subt},
    $[[Γ ⊢ σ2 ○ σ1 ≈ id :  fv iP]]$.
    By a symmetric argument, we also have:
    $[[Γ ⊢ σ1 ○ σ2 ≈ id :  fv iQ]]$.

    Now, we prove that
    $[[Γ ⊢ σ2 ○ σ1 ≈ id :  fv iP]]$ and
    $[[Γ ⊢ σ1 ○ σ2 ≈ id :  fv iQ]]$
    implies that $[[σ1]]$ and $[[σ1]]$
    are (equivalent to) mutually inverse bijections.

@@

\begin{lemma}[Equivalence of polymorphic types]
  \label{lemma:poly-types-equivalence}
  \hfill
  \begin{itemize}
    \item[$-$] For $[[Γ ⊢ ∀pas.iN]]$ and $[[Γ ⊢ ∀pbs.iM]]$,\\ if $[[Γ ⊢ ∀pas.iN ≈ ∀pbs.iM ]]$
    then there exists a bijection $[[μ : {pbs} ∩ fv iM ↔ {pas} ∩ fv iN]]$
    such that $[[ Γ, pas ⊢ iN ≈ [Sub μ] iM ]]$,
    \item[$+$] For $[[Γ ⊢ ∃nas.iP]]$ and $[[Γ ⊢ ∃nbs.iQ]]$,\\  if $[[Γ ⊢ ∃nas.iP ≈ ∃nbs.iQ ]]$
    then there exists a bijection $[[μ : {nbs} ∩ fv iQ ↔ {nas} ∩ fv iP]]$
    such that $[[ Γ, nbs ⊢ iP ≈ [Sub μ] iQ ]]$.
  \end{itemize}
\end{lemma}
\begin{proof}
    \hfill
  \begin{itemize}
    \item[$-$]
      First, by $\alpha$-conversion, we ensure $[[{pas} ∩ fv iM = ∅]]$ and $[[{pbs} ∩ fv iN = ∅]]$.
      By inversion, $[[Γ ⊢ ∀pas.iN ≈ ∀pbs.iM ]]$ implies 
      \begin{enumerate} 
        \item $[[Γ,pbs ⊢ [σ1]iN ≤ iM]]$ for $[[ Γ,pbs ⊢ σ1 :{pas} ]]$ and 
        \item $[[Γ,pas ⊢ [σ2]iM ≤ iN]]$ for $[[ Γ,pas ⊢ σ2 : {pbs} ]]$.
      \end{enumerate}
      To apply \cref{lemma:mutual-subst-subtyping}, we weaken 
      and rearrange the contexts, and extend the substitutions to act as identity
      outside of their initial domain:
      \begin{enumerate} 
        \item $[[Γ, {pas}, {pbs} ⊢ [σ1]iN ≤ iM]]$ for $[[ Γ, {pas}, {pbs} ⊢ σ1 : Γ, {pas}, {pbs} ]]$ and 
        \item $[[Γ, {pas}, {pbs} ⊢ [σ2]iM ≤ iN]]$ for $[[ Γ, {pas}, {pbs} ⊢ σ2 : Γ, {pas}, {pbs} ]]$.
      \end{enumerate}
      Then from \cref{lemma:mutual-subst-subtyping}, 
      there exists a bijection $[[μ : fv iM ↔ fv iN]]$ such that 
      $[[Γ, {pas}, {pbs} ⊢ σ2 ≈ Sub μ :  fv iM]]$ and 
      $[[Γ, {pas}, {pbs} ⊢ σ1 ≈ Sub μ-1 :  fv iN]]$.
